Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-06 doc/manuscript/manuscript.Rmd
update aims
I updated the specific aims and goals of the manuscript. These are
not different than the previous aims, they are just reworded to make
a bit more sense.

We analyzed *S. sclerotiorum* isolates collected from dry bean cultivars from 2003 to 2012 from different geographic locations in America, Europe and Australia using microsatellite markers to delineate genetic differences and reproduction behavior.
The specific aims of our study were:  a) To assess association of phenotypic data (aggressiveness) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S. sclerotiorum* isolates, b) To test the hypothesis that *S. sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S. sclerotiorum* populations are largely clonal, few compare populations from different continents [@lehner2017sclerotinia], c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations. d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.

The overall goal of this work was to identify representative isolates for use in breeding programs.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
Because sampling depended on disease incidence, the number of samples isolated varied from year to year.

@@

Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier [@kohn1990mycelial].
Aggressiveness of each isolate was assessed using a straw test that rates necrotic lesion on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

@@

@@

All data processing and analyses were performed in R version 3.4.0 [@R].
Of the 16 microsatellite loci, five included compound repeats, making allele scoring difficult. 
These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
The clone-corrected data set included 318 isolates.
All analyses were performed on this data set unless otherwise mentioned.
All data and analyses are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/

>>>>

In this study, we analyze and characterize the genetic diversity of 366 *S. sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia. 
Because the white mold nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S. sclerotiorum* populations from white mold nurseries were representative of the fields within the same region. 
As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown among all white mold nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
In total, we collected 366 isolates.

@@

Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
These isolates were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier [@kohn1990mycelial].
Aggressiveness of each isolate was assessed using a straw test that rates necrotic lesion on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

@@

> ZNK: how were the data received and formatted?

@@

Of the 16 microsatellite loci, five included compound repeats, which made allele scoring difficult. 
These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
The clone-corrected data set included 318 isolates.
All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/
